When does vapor pressure deficit drive or reduce evapotranspiration?

### 2018-05-25 doc/paper/body.tex
refactored tex seperate body, removed unknown external dependency on my util package in python modules, and set the stage for removal of tier 2 sites and addition of citation table <credit Trevor Keenan for catch>

\documentclass[12pt]{article}
\input{def.tex}

\begin{document}

\title{When does vapor pressure deficit drive or reduce
  evapotranspiration?}

\author[1]{Adam Massmann\thanks{Corresponding author: akm2203@columbia.edu}}
\author[1]{Pierre Gentine}
\author[1,2]{Changjie Lin}

\affil[1]{Department of Earth and Environmental Engineering, Columbia University, New York, NY 10027}
\affil[2]{State Key Laboratory of Hydroscience and Engineering, Department of Hydraulic
  Engineering, Tsinghua University, Beijing, CN 100084}

\maketitle

\begin{abstract}
  Increasing vapor pressure deficit increases atmospheric demand for water, and vapor pressure deficit is expected to rise with increasing greenhouse gases. While increased evapotranspiration (ET) in response to increased atmospheric demand seems intuitive, plants are capable of reducing ET in response to increased VPD by closing their stomata, in an effort to conserve water. Here we examine which effect dominates response to increasing VPD: atmospheric demand and increases in ET, or plant physiological response (stomata closure) and decreases in ET. We use Penman-Monteith, combined with semi-empirical optimal stomatal regulation theory and underlying water use efficiency, to develop a theoretical framework for understanding how ET responds to increases in VPD.

 The theory suggests that for most environmental conditions and plant types, plant physiological response dominates and ET decreases with increasing VPD. Plants that are evolved or bred to prioritize primary production over water conservation (e.g. crops) exhibit a higher likelihood of atmospheric demand-driven response (ET increasing). However for forest, grass, savannah, and shrub plant types, ET more frequently decreases than increases with rising VPD. This work serves as an example of the utility of our simplified framework for disentangling land-atmosphere feedbacks at multiple scales, including the characterization of ET response in an atmospherically drier, enriched CO$_2$ world. 
\end{abstract}

\onehalfspacing

@@

\section{Materials and Methods}
\subsection{Data}
\label{data}
We use both meteorological and eddy-covariance data from the
FLUXNET2015 database (data available at \sloppy
\url{https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/} \sloppy),
including all sites with at least four years of data and observations
of the variables described in the methods section (Section
\ref{methods}). Seventy-three sites met these requirements, and were
grouped into nine plant functional types (PFT) according to the
International Geosphere-Biosphere Programme vegetation classification
scheme \citep{Loveland_1999}: cropland (CRO), grass (GRA), deciduous
broadleaf forest (DBF), evergreen broadleaf forest (EBF), evergreen
needleleaf forest (ENF), mixed forest (MF), closed shrub (CSH),
savannah (SAV), and woody savannah (WSA) (site locations in Figure
\ref{map_fig}).

@@

@@

\acknowledgments This work used eddy covariance data acquired and
shared by the FLUXNET community, including these networks: AmeriFlux,
AfriFlux, AsiaFlknux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont,
ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC,
OzFlux-TERN, TCOS-Siberia, and USCCC. The ERA-Interim reanalysis data
are provided by ECMWF and processed by LSCE. The FLUXNET eddy
covariance data processing and harmonization was carried out by the
European Fluxes Database Cluster, AmeriFlux Management Project, and
Fluxdata project of FLUXNET, with the support of CDIAC and ICOS
Ecosystem Thematic Center, and the OzFlux, ChinaFlux and AsiaFlux
offices. This material is based upon work supported by the National
Science Foundation Graduate Research Fellowship under Grant No. DGE
16-44869. All code and data used in this analysis, including those
used to generate the figures and tables, are publicly available at
\url{https://github.com/massma/climate\_et}.
%\hyperref[https://github.com/massma/climate\_et]{https://github.com/massma/climate_et}.

>>>>

@@

\section{Materials and Methods}
\subsection{Data}
\label{data}
We use both meteorological and eddy-covariance data from the
FLUXNET2015 database (data available at \sloppy
\url{https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/} \sloppy),
including all sites with at least four years of data and observations
of the variables described in the methods section (Section
\ref{methods}). Seventy-three sites met these requirements, and were
grouped into nine plant functional types (PFT) according to the
International Geosphere-Biosphere Programme vegetation classification
scheme \citep{Loveland_1999}: cropland (CRO), grass (GRA), deciduous
broadleaf forest (DBF), evergreen broadleaf forest (EBF), evergreen
needleleaf forest (ENF), mixed forest (MF), closed shrub (CSH),
savannah (SAV), and woody savannah (WSA) (site locations and citations in Figure
\ref{map_fig} and Table \ref{flux_sites}).

@@

\begin{table}
  \caption{Metadata and citations for flux sites used in this analysis. All data are gathered from \url{www.fluxdata.org} using tools available at \url{https://github.com/trevorkeenan/FLUXNET_citations}.}
  \label{flux_sites}
  \label{definitions}
  \centering \footnotesize
  \begin{tabular}{l l l l l l l}
    \hline
    Site name & PFT & Lat & Lon & Clim$^{a}$ & Period & References \\
    \hline
    \input{flux_sites.tex}
    \hline
    \multicolumn{7}{l}{$^{a}$K{\"o}ppen Climate classification.}
    \hline
\end{table}

@@

\acknowledgments This work used eddy covariance data acquired and shared
by the FLUXNET community, including these networks: AmeriFlux, AfriFlux,
AsiaFlknux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont,
ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC,
OzFlux-TERN, TCOS-Siberia, and USCCC. The ERA-Interim reanalysis data
are provided by ECMWF and processed by LSCE. The FLUXNET eddy covariance
data processing and harmonization was carried out by the European Fluxes
Database Cluster, AmeriFlux Management Project, and Fluxdata project of
FLUXNET, with the support of CDIAC and ICOS Ecosystem Thematic Center,
and the OzFlux, ChinaFlux and AsiaFlux offices. We would like to think
Dr. Trevor Keenan for providing tools for citation of the FLUXNET2015
dataset, available at
\url{https://github.com/trevorkeenan/FLUXNET_citations}. This material is
based upon work supported by the National Science Foundation Graduate
Research Fellowship under Grant No. DGE 16-44869. All code and data used
in this analysis, including those used to generate the figures and
tables, are publicly available at
\url{https://github.com/massma/climate\_et}.
%\hyperref[https://github.com/massma/climate\_et]{https://github.com/massma/climate_et}.
